### 2020-02-05 ncov.tex
final update before I go to bed

As the disease continues to spread, many researchers have already published their analyses of the outbreak as pre-prints (e.g., \cite{bedfordncov, imaincov, liuncov, majumderncov, readncov, zhaoncov}) and in peer-reviewed journals (e.g., \cite{li2020early, riou2020pattern, wu2020nowcasting}), focusing in particular on estimates of the basic reproductive number \Ro (i.e., the average number of secondary cases generated by a primary case in a fully susceptible population \citep{anderson1991infectious, diekmann1990definition}).
Estimates of the basic reproductive number are of interest during an outbreak because they provide information about the level of intervention required to interrupt transmission \citep{anderson1991infectious}, and about the potential final size of the outbreak \citep{anderson1991infectious, ma2006generality}.
We commend these researchers for their timely contribution and those who made the data publicly available.
However, it can be difficult to compare a disparate set of estimates of \Ro from different research groups (as well as the associated degrees of uncertainty) when the estimation methods and their underlying assumptions vary widely.

@@

We reviewed 7 modeling studies of the novel coronavirus outbreak that were published online between January 23 and January 26, 2020 (Table 1).
Five studies \citep{liuncov, majumderncov, readncov, riouncov, zhaoncov} were uploaded to pre-print servers (bioRxiv, medRxiv, and SSRN); one report was posted on the website of Imperial College London \citep{imaincov}; and one report was posted on nextstrain.org \citep{bedfordncov}.
They vary widely in their statistical methods and in the amount of data used to infer \Ro.
\cite{bedfordncov}, \cite{imaincov}, and \cite{riouncov} simulated branching process models and compared the predicted number of cases from their models with the estimated number of total cases by January 18 (between 1,000 and 9,7000 \citep{imaincov0}).
\cite{readncov} fitted a deterministic, metapopulation Susceptible-Exposed-Infected-Recovered (SEIR) model to incidence data between January 1 and January 21 from major cities in China and other countries.
\cite{zhaoncov} and \cite{liuncov} fitted exponential growth models to incidence data up to January 22 and January 23, respectively, and inferred \Ro via the Euler-Lotka equation (\eref{euler}).
\cite{majumderncov} fitted the Incidence Decay and Exponential Adjustment (IDEA) model \citep{fisman2013idea} to incidence data up to January 26, which is equivalent to fitting an exponential growth model and assuming a fixed generation-interval distribution.

>>>>

As the disease continues to spread, many researchers have already published their analyses of the outbreak as pre-prints (e.g., \cite{bedfordncov, imaincov, liuncov, majumderncov, readncov, zhaoncov}) and in peer-reviewed journals (e.g., \cite{li2020early, riou2020pattern, wu2020nowcasting, zhao2020preliminary}), focusing in particular on estimates of the basic reproductive number \Ro (i.e., the average number of secondary cases generated by a primary case in a fully susceptible population \citep{anderson1991infectious, diekmann1990definition}).
Estimates of the basic reproductive number are of interest during an outbreak because they provide information about the level of intervention required to interrupt transmission \citep{anderson1991infectious}, and about the potential final size of the outbreak \citep{anderson1991infectious, ma2006generality}.
We commend these researchers for their timely contribution and those who made the data publicly available.
However, it can be difficult to compare a disparate set of estimates of \Ro from different research groups (as well as the associated degrees of uncertainty) when the estimation methods and their underlying assumptions vary widely.

@@

We reviewed 7 modeling studies of the novel coronavirus outbreak that were published online between January 23 and January 26, 2020 (Table 1).
Five studies \citep{liuncov, majumderncov, readncov, riouncov, zhaoncov} were uploaded to pre-print servers (bioRxiv, medRxiv, and SSRN); one report was posted on the website of Imperial College London \citep{imaincov}; and one report was posted on nextstrain.org \citep{bedfordncov}.
They vary widely in their statistical methods and in the amount of data used to infer \Ro.
\cite{bedfordncov}, \cite{imaincov}, and \cite{riouncov} simulated branching process models and compared the predicted number of cases from their models with the estimated number of total cases by January 18 (between 1,000 and 9,7000 \citep{imaincov0}).
\cite{readncov} fitted a deterministic, metapopulation Susceptible-Exposed-Infected-Recovered (SEIR) model to incidence data between January 1 and January 21 from major cities in China and other countries.
\cite{zhaoncov} and \cite{liuncov} fitted exponential growth models to incidence data up to January 22 and January 23, respectively, and inferred \Ro via the Euler-Lotka equation (\eref{euler}).
\cite{majumderncov} fitted the Incidence Decay and Exponential Adjustment (IDEA) model \citep{fisman2013idea} to incidence data up to January 26, which is equivalent to fitting an exponential growth model and assuming a fixed generation-interval distribution.
Some of these studies have now been published in peer-reviewed journals \citep{riou2020pattern, zhao2020preliminary} or have been updated with better uncertainty quantification \citep{readncov2}.
\swp{I think it's fair to add the last sentence because we do not want to misrepresent these studies.}
